---
title: "Эконометрикс II"
subtitle: "Лекц 6: Мөчлөгийг тодорхойлох<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "багш: Д. Хашбаатар"
date: "2019/09/23"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    nature: 
      beforeInit: "macros.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Улиралын нөлөөг загварчлах

Хугацаан цуваа
```{r}
library(kableExtra)
kable(cbind(mtcars, mtcars)) %>%
  kable_styling() %>%
  scroll_box(width = "500px", height = "200px")
```

---
#White noise

```{r}
set.seed(1)
plot(rnorm(100), type="l", ylab="Process", main="Realization of White Noise Process")
```

---
<div class = "keyconcept" id="KC4.1">
<h3 class = "right"> Key Concept 4.1 </h3>          
<h3 class = "left"> Terminology for the Linear Regression Model with a Single Regressor </h3>

<p> The linear regression model is 

$$Y_i = \beta_0 + \beta_1 X_i + u_i$$

where

- the index $i$ runs over the observations, $i=1,\dots,n$
- $Y_i$ is the *dependent variable*, the *regressand*, or simply the *left-hand variable*
- $X_i$ is the *independent variable*, the *regressor*, or simply the *right-hand variable*
- $Y = \beta_0 + \beta_1 X$ is the *population regression line* also called the *population regression function*
- $\beta_0$ is the *intercept* of the population regression line
- $\beta_1$ is the *slope* of the population regression line
- $u_i$ is the *error term*.
</p>
</div>